Naive Bayes Workflow Ver 2.0.yn

### 2019-01-31 Naive_bayes_V2yn.Rmd
Update

##Summary
This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

@@

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)

#Rename the columns.
names(comments_rawVERB)[1] <- "subject"
names(comments_rawVERB)[2] <- "structure"
names(comments_rawVERB)[3] <- "text"

#Change "subject" element from character to a factor for analysis.
comments_rawVERB$subject <- factor(comments_rawVERB$subject)
str(comments_rawVERB$subject)
table(comments_rawVERB$subject)
```

```{r}
#Change "structure" element from character to a factor for analysis.
comments_rawVERB$structure <- factor(comments_rawVERB$structure)
str(comments_rawVERB$structure)
table(comments_rawVERB$structure)
```

@@

```{r}
#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
comments_rawVERB$randomizer <- c(vector)
comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
print(comments_rawVERB_R)

@@

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB_R[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB_R[[5]])
```

@@

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)

#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())

@@

#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

@@

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

@@

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
```

@@

>>>>

##Summary
This is **Version 2.0.yn** of a basic workflow for classifying TA comments using a naive Bayes model. It is optimized for testing pair-wise assignment to categories (Yes/No decision) beginning with the most frequently encountered:

*  3_technical for the Subject category
*  4_specific for the Structure category

The main features and permutations to be evaluated are:

*  Stepwise paired classification versus comparing multiple groups simultaneously

Additional working notes are located at the end of this page.

@@

##Initial Setup
Calling required libraries.

```{r}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

@@

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVerYN.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVerYN <- filter(base_dataVerYN,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVerYN <- comments_subsetVerYN %>% select(23,24,22)

#Rename the columns.
names(comments_rawVerYN)[1] <- "subject"
names(comments_rawVerYN)[2] <- "structure"
names(comments_rawVerYN)[3] <- "text"

#Change "subject" element from character to a factor for analysis.
comments_rawVerYN$subject <- factor(comments_rawVerYN$subject)
str(comments_rawVerYN$subject)
table(comments_rawVerYN$subject)
```

```{r}
#Change "structure" element from character to a factor for analysis.
comments_rawVerYN$structure <- factor(comments_rawVerYN$structure)
str(comments_rawVerYN$structure)
table(comments_rawVerYN$structure)
```

@@

```{r}
#Append the vector as a new column to "comments_rawVerYN" with name "randomizer"
#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
comments_rawVerYN$randomizer <- c(vector)
comments_rawVerYN_R <-comments_rawVerYN %>% arrange(randomizer)
print(comments_rawVerYN_R)

@@

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVerYN_R <- VCorpus(VectorSource(comments_rawVerYN_R$text))
print(comments_corpusVerYN_R)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVerYN_R[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVerYN_R[[5]])
```

@@

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVerYN_R <- tm_map(comments_corpusVerYN_R, content_transformer(tolower))

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removePunctuation)

#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?
comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removeNumbers)

#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it
comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removeWords, stopwords())

@@

#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVerYN_R[[5]]))
```

@@

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVerYN_R, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

@@

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVerYN_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVerYN_R[7006:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVerYN_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVerYN_R[7006:9340,]$structure
```

@@

##Version 2.0.yn of Naive Bayes (NB) Protocol
###Initial Data Input
```{}
#Read in TA comments from CSV file.
base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVerYN.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{}
#Select rows representing the sub-groups to compare. 
comments_subsetVerYN <- filter(base_dataVerYN,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVerYN <- comments_subsetVerYN %>% select(23,24,22)

#Rename the columns.
names(comments_rawVerYN)[1] <- "type"
names(comments_rawVerYN)[2] <- "locus"
names(comments_rawVerYN)[3] <- "text"

#Change "type" element from character to a factor for analysis.
comments_rawVerYN$type <- factor(comments_rawVerYN$type)
str(comments_rawVerYN$type)
table(comments_rawVerYN$type)
```

```{}
#Change "locus" element from character to a factor for analysis.
comments_rawVerYN$locus <- factor(comments_rawVerYN$locus)
str(comments_rawVerYN$locus)
table(comments_rawVerYN$locus)
```

***
***
Here is where I need to add the column used to randomly assign the data to training or testing sets.

Use set.seed(1234) to create a stable randomized set.

***
***

Convert data set to a volative corpus using "tm" library.
```{}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVerYN <- VCorpus(VectorSource(comments_rawVerYN$text))
print(comments_corpusVerYN)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVerYN[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVerYN[[5]])
```

```{}
#Convert text to all lower case letters.
comments_corpus_cleanVerYN <- tm_map(comments_corpusVerYN, content_transformer(tolower))
```

```{}
#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, removePunctuation)
```

```{}
#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?

comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, removeNumbers)
```

```{}
#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it

comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, removeWords, stopwords())
```

Option to stem terms has been removed from this version.

Final prep and data check.
```{}
#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVerYN[[5]]))
```

```{}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVerYN, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
```{}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{}
comments_train_labels_type <- comments_rawVerYN[1:7005, ]$type
comments_test_labels_type <- comments_rawVerYN[7006:9340,]$type
```

```{}
comments_train_labels_locus <- comments_rawVerYN[1:7005, ]$locus
comments_test_labels_locus <- comments_rawVerYN[7006:9340,]$locus
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{}
prop.table(table(comments_train_labels_type))
prop.table(table(comments_test_labels_type))
```

```{}
prop.table(table(comments_train_labels_locus))
prop.table(table(comments_test_labels_locus))
```

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no.

```{}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

###Train Model, Predict, Evaluate for Type
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{}
comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
comments_test_pred_type <- predict(comments_classifier_type, comments_test)
CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.

```{}
print("Overall TYPE accuracy")
(14+45+1346+19)/23.35
```

###Train Model, Predict, Evaluate for LOCUS
This is the same block as above but uses the "Locus" data column.

```{}
comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.

```{}
print("Overall LOCUS accuracy")
(62+64+931+3)/23.35
```

```{}
predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
head(predicted_category_type)
write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
```

```{}
predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
head(predicted_category_locus)
write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
```

***
START COMMENTS
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.

Permutations to Test:

*  Compare 1-grams, 2-grams, 3-grams
*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
    +  Initial approach will be to split between largest group and balance in each category
    +  Ex.: type is 3-technical vs. not 3_technical

Known Issues:

1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
